---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# lmtp

<!-- badges: start -->
[![Build Status](https://travis-ci.com/nt-williams/lmtp.svg?token=DA4a53nWMx6q9LisKdRD&branch=master)](https://travis-ci.com/nt-williams/lmtp)
[![codecov](https://codecov.io/gh/nt-williams/lmtp/branch/master/graph/badge.svg?token=TFQNTischL)](https://codecov.io/gh/nt-williams/lmtp)
[![R build status](https://github.com/nt-williams/lmtp/workflows/R-CMD-check/badge.svg)](https://github.com/nt-williams/lmtp/actions)
<!-- badges: end -->

> Non-parametric Causal Effects Based on Longitudinal Modified Treatment Policies

# Installation

The development version can be installed from GitHub with: 

```r
devtools::install_github("nt-williams/lmtp")
```

# Example

We have a data set with treatment nodes at 4 time points and a binary outcome at time 5.

```{r}
head(lmtp::sim_t4)
```

We're interested in the effect of an intervention on outcome `Y` where treatment is decreased by 1 at all time points for observations whose treatment value won't go below 0 if intervened upon. The true value under this intervention is about 0.48.

![](https://gist.githubusercontent.com/nt-williams/2488fef9e94c7ef1a3920c2682433980/raw/2a339c10d651aa18dc48188b4017c605c30e2405/lmtp-readme-example.svg?sanitize=true)


